Whitepaper: Contract Modules

### 2020-11-26 papers/contract-modules/modules.tex
even more specific - fragments of an implementation

\lstset{
  language=caml,
  morekeywords={contract,sig,entrypoint,module,raises,val},
  % predefined types
  classoffset=1,
  morekeywords={nat,int,unit,string},
  keywordstyle=\color{blue},
  classoffset=2,
  morekeywords={Left,Right},
  keywordstyle=\color{magenta},
  backgroundcolor=\color{brown!20}
}

@@

\title{Contract Modules}
\author{Peter Thiemann}
\institute{}

@@

Here is a contract type that describes / specifies the above Michelson code.
\begin{lstlisting}[caption={Contract module example},label={lst:contract-module-example}]
contract type MyContract = sig
  entrypoint %bid_for_auction (_ : int) 
  raises "closed" as closed       (** auction closed*)
       | "bid too low" as too_low (** bid received is too low *)
       | 42 as the_answer         (** failwith received argument 42 *)
       | (_ : int) as failedwith  (** failwith received an integer arg *)

@@

@@

\begin{lstlisting}
...
  entrypoint bid_for_auction (Left (_ : int))
...
  entrypoint close_auction (Right (_ : unit))
\end{lstlisting}

@@

\begin{lstlisting}
module type MyContract =
sig
  type bid_for_auction_errors = 
     | closed       (** auction closed*)
     | too_low      (** bid received is too low *)
     | the_answer   (** failwith received argument 42 *)
     | failedwith of int (** failwith received an integer arg *)

  val bid_for_auction
    : int -> (status, bid_for_auction_errors) Tezos.monad

@@

  val close_auction
    : unit -> (status, close_auction_errors) Tezos.monad
end
\end{lstlisting}

\begin{itemize}
\item The type \lstinline/Tezos.monad/ is defined by the underlying
  API. The untyped API uses a type \lstinline/Answer.t/. The exact
  relationship is TBD.
\item Error reporting assumes that the monad returns a sum type. But
  things could be handled differently (for the
  \lstinline/close_auction/entrypoint): by an error continuation of
  type \lstinline/close_auction_errors -> Answer.t/ or by
  raising an exception that encapsulates the error type:
\begin{lstlisting}
exception Close_Auction_Exn of close_auction_errors
\end{lstlisting}
\end{itemize}

@@

If the contract source is unavailable, then we can obtain it from the
contract's public hash using an API call. In that case, we would
specify the hash in the contract header as follows
\begin{lstlisting}
contract type MyContract "KT1ThEdxfUcWUwqsdergy..." = sig
...
\end{lstlisting}

We propose to deal with the remaining errors (i.e., built-in errors
and failures that are not specifed) like Java's unchecked
exceptions. That is, the API function throws an exception.

@@

@@

>>>>

\lstset{
  language=caml,
  morekeywords={contract,entrypoint,module,hash,paid,raises,sig,val},
  % predefined types
  classoffset=1,
  morekeywords={nat,int,unit,string,pukh,status,tz},
  keywordstyle=\color{blue},
  classoffset=2,
  morekeywords={Left,Right},
  keywordstyle=\color{magenta},
  backgroundcolor=\color{brown!20},
  captionpos=b,
  literate={->}{{$\to$}}1
}

@@

\title{Whitepaper: Contract Modules}
\author{Peter Thiemann}
\institute{}

@@

Here is a contract type that describes / specifies the above Michelson code.
\begin{lstlisting}[caption={Contract module example},label={lst:contract-module-example}]
contract type MyContract = sig
  paid entrypoint %bid_for_auction (value : int) 
  raises "closed"                 (** auction closed*)
       | "bid too low" as too_low (** bid received is too low *)
       | 42 as the_answer         (** failwith received argument 42 *)
       | (_ : int) as failedwith  (** failwith received an integer arg *)

@@

The body of the \lstinline/sig ... end/ can be stored in a
\texttt{.tzm} file where the basename matches the one of the
\texttt{.tz} file. The output is a \texttt{.mli} file which contains a
valid OCaml signature. In the next step, we are also generating the
corresponding \texttt{.ml} file implementing the signature.

Each \lstinline/entrypoint/ is mapped to a function. The final pattern
variable may be a wildcard \lstinline/_/ or an OCaml identifier, which
is used in generating the implementation. For \lstinline/_/ a standard
name will be generated.

A \lstinline/paid/ entrypoint requires the caller to specify an amount of tokens to
accompany the contract invocation. Accordingly, the generated function has an extra
argument. Otherwise, the entrypoint is unpaid.

The \lstinline/raises/ clauses specify values that the contract might
pass to the \texttt{FAILWITH} instruction. The \lstinline/as/ clause
maps the matched value to an OCaml identifier, which is used to
generate the corresponding case in the error type. There is a standard
translation (to be specified) that maps a value without an
\lstinline/as/ clause to an identifier.

@@

\begin{lstlisting}[caption={Alternative syntax for entrypoints},label={lst:alternative-syntax}]
  paid entrypoint bid_for_auction (Left (_ : int))
...
  entrypoint close_auction (Right (_ : unit))
\end{lstlisting}

@@

\begin{lstlisting}[caption={Generated signature},label={lst:generated-signature}]
module type MyContract =
sig
  type bid_for_auction_errors = 
     | closed       (** auction closed*)
     | too_low      (** bid received is too low *)
     | the_answer   (** failwith received argument 42 *)
     | failedwith of int (** failwith received an integer arg *)

  val bid_for_auction
    : Tezos.pukh -> ?fee:Tezos.tz -> Tezos.tz
    -> int -> (Tezos.status, bid_for_auction_errors) Tezos.monad

@@

  val close_auction
    : Tezos.pukh -> ?fee:Tezos.tz -> ?amount:Tezos.tz
    -> unit -> (Tezos.status, close_auction_errors) Tezos.monad
end
\end{lstlisting}

\begin{itemize}
\item The first argument is the public key hash of the caller. 
\item The \lstinline/paid/ entrypoint requires an \lstinline/amount/
  argument, which is optional (with default 0) for an unpaid
  entrypoint.
\item The \lstinline/fee/ parameter is optional in both
  cases. It is precomputed and supplied as the default value in the
  implementation.

  Q (extension): Would it make sense to specify something about fees in the
  contract module? For example, we might specify a fee schedule which
  raises the fee until the call is accepted.
\item The type \lstinline/Tezos.monad/ is defined by the underlying
  API. The untyped API uses a type \lstinline/Answer.t/. The exact
  relationship is TBD.
\item Error reporting assumes that the monad returns a sum type. But
  things could be handled differently (for the
  \lstinline/close_auction/entrypoint): by an error continuation of
  type \lstinline/close_auction_errors -> Answer.t/ or by
  raising an exception that encapsulates the error type:
\begin{lstlisting}
exception Close_Auction_Exn of close_auction_errors
\end{lstlisting}
\end{itemize}

@@

If the contract source is unavailable, then we can obtain it from the
contract's public hash using an API call. In that case, we would
specify the hash in the contract.
\begin{lstlisting}
sig
  hash "KT1ThEdxfUcWUwqsdergy..."
...
\end{lstlisting}

We propose to treat the remaining errors (i.e., built-in errors
and failures that are not specifed) like Java's unchecked
exceptions. That is, the API function throws an exception.

\begin{lstlisting}[caption={Generated implementation},label={lst:generated-implementation}]
(* constant template *)
open Tezos_api
open Tezos_serialization (* Tint, Tunit, serialize, ... *)

let failwith_regex = 
  "A FAILWITH instruction was reached.*"
let int_regex = 
  "[+-]?[0-9]+"

(* variable part *)
let contract_addr = 
  get_contract "KT1ThEdxfUcWUwqsdergy..." >>=? fun contract -> ???

type bid_for_auction_errors = 
   | closed       (** auction closed*)
   | too_low      (** bid received is too low *)
   | the_answer   (** failwith received argument 42 *)
   | failedwith of int (** failwith received an integer arg *)

let big_for_auction_errors_map =
  let open Base in
  Base.Map.of_alist_exn (module String)
    [
      (failwith_regex ^ "closed", closed);
      (failwith_regex ^ "bid too low", too_low);
      (failwith_regex ^ "42", the_answer);
      (failwith_regex ^ int_regex, failedwith (0)); (* badness 10000 *)
    ]
  (* inadequate in several ways;
   * should just generate code instead of using a map*)

let bid_for_auction destination ?(fee=100.0) amount value =
  ...
  begin
    SyncAPIV0.call_contract
      amount
      destination
      contract_addr
      ~(serialize Tint value)
      fee
    >>=? fun oph ->
    wait_for_inclusion oph (* busy waiting *)
    >>=? fun _ ->
    return 0
  end
  >>= function
  | Ok _ -> ...
  | Error (Rejection Michelson_runtime_error s) -> ...

let close_auction destination ?(fee=50.0) ?(amount=0.0) _ =
  ...
  SyncAPIV0.call_contract
    amount
    destination
    contract_addr
    ~(serialize Tunit)
    fee
\end{lstlisting}

@@

The checker should warn about unhandled failures (both checked and unchecked).

@@

The checker should also make sure that unpaid entrypoints do not
depend on a non-zero amount being passed to the contract
invocation. For example, the \texttt{AMOUNT} instruction of Michelson
should not be used in the code.

The checker should precalculate the fee. It essentially obtains
this value from the Tezos library.

Failures might arise in contracts that are transitively invoked. In
this case, the checker would obtain the callees' code and process
it. The first step, not too difficult, in this direction would be to
approximate the contents of the list of operations that is produced by
the contract.
